Supporting Information for\\ ``A heterogeneous landscape of selection and epistasis in protein-coding genes''

### 2022-03-27 manuscript/supplement.tex
finish up supplement

Simulations had diploid population sizes of 5000 and I drew haploid sample
sizes of 50 (25 diploids), with recombination rates varying from \(0\) to
\(0.001\) (or \(\rho=0-10\)). I tested selection coefficients of \(s\) of 0,
\(-2\times10^{-5}\), \(-2\times10^{-4}\), and \(-2\times10^{-3}\) (or
\(\gamma=0\), \(-0.2\), \(-2\), and \(-10\)). Epistasis was set to
\(\epsilon=0\), \(-0.5\), and \(0.5\). Conditional distributions of the
two-locus sampling distribution show good agreement between numerical solutions
and simulations, and residuals do not show any directional bias
(Figures~\ref{fig:validation}--\ref{fig:last_validation}).

@@

\begin{itemize}
    \item Why did we do it
    \item How did we do it.. quite simple
\end{itemize}

@@

\subsection{Multinucleotide mutations and positive LD between linked synonymous variants}

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{llll}
        \toprule
        Diploid genotype & General model & Simple dominance & Gene-based dominance\\
        \midrule
        $AB$ / $AB$ & $1 + s_{AB/AB}$ & $1 + 2s_A + 2s_B$ & $1 + 2s$\\
        $AB$ / $Ab$ & $1 + s_{AB/Ab}$ & $1 + 2s_A + 2s_Bh_B$ & $1 + 2s$\\
        $AB$ / $aB$ & $1 + s_{AB/aB}$ & $1 + 2s_Ah_A + 2s_B$ & $1 + 2s$\\
        $AB$ / $ab$ & $1 + s_{AB/ab}$ & $1 + 2s_Ah_A + 2s_Bh_B$ & $1 + 2sh$\\
        $Ab$ / $Ab$ & $1 + s_{Ab/Ab}$ & $1 + 2s_A$ & $1 + 2s$\\
        \addlinespace
        $Ab$ / $aB$ & $1 + s_{Ab/aB}$ & $1 + 2s_Ah_A + 2s_Bh_B$ & $1 + 2s$\\
        $Ab$ / $ab$ & $1 + s_{Ab/ab}$ & $1 + 2s_Ah_A$ & $1 + 2sh$\\
        $aB$ / $aB$ & $1 + s_{aB/aB}$ & $1 + 2s_B$ & $1 + 2s$\\
        $aB$ / $ab$ & $1 + s_{aB/ab}$ & $1 + 2s_Bh_B$ & $1 + 2sh$\\
        $ab$ / $ab$ & $1$ & $1$ & $1$\\
        \bottomrule
    \end{tabular}
    \caption{General selection model for diploids and dominance models.}
    \label{tab:selmodels}
\end{table}

\begin{table}[ht!]
    \centering
    \begin{tabular}{ll}
        \toprule
        Haplotype & Fitness\\
        \midrule
        $AB$ & $(1 + s_A + s_B)(1+\epsilon)$\\
        $Ab$ & $1 + s_A$\\
        $aB$ & $1 + s_B$\\
        $ab$ & $1$\\
        \bottomrule
    \end{tabular}
    \caption{Haploid epistasis model.}
    \label{tab:epistasis}
\end{table}

\begin{table}[ht!]
    \centering
    \begin{tabular}{lll}
        \toprule
        Code & Description & Region\\
        \midrule
        ESN & Esan in Nigeria & Africa\\
        GWD & Gambian in Western Divisions in the Gambia & Africa\\
        LWK & Luhya in Webuye, Kenya & Africa\\
        MSL & Mende in Sierra Leone & Africa\\
        YRI & Yoruba in Ibadan, Nigeria & Africa\\
        \addlinespace
        CEU & Utah Residents (CEPH) with Northern and Western European Ancestry & Europe\\
        GBR & British in England and Scotland & Europe\\
        FIN & Finnish in Finland & Europe\\
        IBS & Iberian Population in Spain & Europe\\
        TSI & Toscani in Italia & Europe\\
        \addlinespace
        CDX & Chinese Dai in Xishuangbanna, China & East Asia\\
        CHB & Han Chinese in Beijing, China & East Asia\\
        CHS & Southern Han Chinese & East Asia\\
        JPT & Japanese in Tokyo, Japan & East Asia\\
        KHV & Kinh in Ho Chi Minh City, Vietnam & East Asia\\
        \bottomrule
    \end{tabular}
    \caption{Thousand Genomes Project population descriptions for populations used in this study.}
    \label{tab:1kgpops}
\end{table}

\begin{longtable}[t]{lllr}
\caption{\label{tab:tajimasD}Tamija's $D$ for classes of coding mutations, both within annotated domains and outside of domains.}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endfirsthead
\caption[]{\label{tab:tajimasD}Tamija's $D$ for classes of coding mutations, both within annotated domains and outside of domains. \textit{(continued)}}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endhead

@@

\begin{table}
    \centering
    \small
    \begin{tabular}{lrrrrrrrrr}
        \toprule
        Class & $h$ & shape & scale & LL &
        $[0,10^{-5})$ & $[10^{-5},10^{-4})$ & $[10^{-4},10^{-3})$ &
        $[10^{-3},10^{-2})$ & $[10^{-2},\infty)$\\
        \midrule
        Missense & 0.0 & 0.093 & 768505 & -678.2 & 0.260 & 0.062 & 0.077 & 0.096 & 0.505\\
         & 0.2 & 0.138 & 6660 & -416.7 & 0.260 & 0.098 & 0.134 & 0.182 & 0.327\\
         & 0.5 & 0.147 & 2117 & -392.0 & 0.282 & 0.114 & 0.159 & 0.214 & 0.231\\
        \addlinespace
        LOF & 0.0 & 0.132 & 99999054 & -248.3 & 0.077 & 0.028 & 0.037 & 0.051 & 0.807\\
         & 0.2 & 0.177 & 477994 & -226.7 & 0.083 & 0.042 & 0.063 & 0.095 & 0.717\\
         & 0.5 & 0.188 & 121419 & -224.2 & 0.092 & 0.050 & 0.077 & 0.119 & 0.662\\
        \bottomrule
    \end{tabular}
    \caption{
        DFEs inferred for missense and loss-of-function variants in MSL for varying
        values of $h$. General patterns are consistent across different chosen values
        of $h$, although $h=0$ results in poorer fits for both missense and LOF variants.
        Columns to the right of the log-likelihood (LL) column show proportions of
        new mutations with $|s|$ in each given bin.
    }
    \label{tab:msldfe}
\end{table}

@@

\textbf{all supporting figures need captions}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/jackknife}
    \caption{
        \textbf{Accuracy of the jackknife approximation and runtime.}
        Small sample sizes can lead to large error in the moment-closure
        approximation for larger recombination distances or selection
        coefficients. Generally, the jackknife approximation breaks down for
        recombination rates greater than \(\rho\approx30\). While increasing
        sample size leads to more accurate solutions, it comes at the cost of
        both increased runtime and memory usage. Most analyses in this paper
        used sample sizes between 40 and 80.
    }
    \label{fig:jackknife}
\end{figure}

\begin{figure}[ht!]
    \centering
    %\includegraphics{}
    \caption{\textbf{Validation figures.}}
    \label{fig:validation}
\end{figure}

\begin{figure}[ht!]
    \centering
    %\includegraphics{}
    \caption{\textbf{Validation figures.}}
    \label{fig:last_validation}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with slightly deleterious mutations.}
        Using an individual based forward simulator
    }
    \label{fig:bgs1}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with moderately deleterious mutations.}
    }
    \label{fig:bgs2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.0_n_50}
    \caption{
        \textbf{Associative overdominance due to slightly deleterious recessive mutations.}
    }
    \label{fig:bgs3}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.1_n_80}
    \caption{
        \textbf{Background selection with moderately deleterious recessive mutations.}
    }
    \label{fig:bgs4}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.sd2}
    \caption{
        \textbf{.}
    }
    \label{fig:toy_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance}
    \caption{
        \textbf{.}
    }
    \label{fig:toy_dom}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance.sd2}
    \caption{
        \textbf{.}
    }
    \label{fig:toy_dom_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU}
    \caption{
        \textbf{.}
    }
    \label{fig:relate}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.sd2}
    \caption{
        \textbf{.}
    }
    \label{fig:relate_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.dominance.sd2}
    \caption{
        \textbf{.}
    }
    \label{fig:relate_dom_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_gene_wide}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        in the same gene.}
    }
    \label{fig:LDgene}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_in_domain}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        that fall inside the same domain.}
    }
    \label{fig:LDwithin}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_outside_domains}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        outside of domains.}
    }
    \label{fig:LDoutside}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_afr}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
    }
    \label{fig:domainsAFR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eur}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
    }
    \label{fig:domainsEUR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eas}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
    }
    \label{fig:domainsEAS}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../figures/data_within_between}
    \caption{
        \textbf{Within vs between.}
    }
    \label{fig:between_domains}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/mutations_in_genes}
    \caption{
        \textbf{Correlation between numbers of missense and synonymous mutations
        within each gene.}
    }
    \label{fig:mutCorrGene}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/mutations_in_domains}
    \caption{
        \textbf{Correlation between numbers of missense and synonymous mutations
        within each annotated domain.}
    }
    \label{fig:mutCorrDomain}
\end{figure}

>>>>

Simulations had diploid population sizes of 5000 and I drew haploid sample
sizes of 50 (25 diploids), with recombination rates varying from \(0\) to
\(0.001\) (or \(\rho=0-20\)). I tested selection coefficients of \(s\) of 0,
\(-1\times10^{-5}\), \(-2\times10^{-4}\), and \(-2\times10^{-3}\) (or
\(\gamma=0\), \(-0.1\), \(-2\), and \(-20\)). Epistasis was set to
\(\epsilon=0\), \(-0.5\), and \(0.5\), and dominance was set to either
\(h=0.5\) (additive) or \(0.1\) (recessive). Conditional distributions of the
two-locus sampling distribution show good agreement between numerical solutions
and simulations, and residuals do not show any directional bias
(Figures~\ref{fig:validation}--\ref{fig:last_validation}). The Support Figures
includes a handful of these comparisons, and the full set of comparisons (over
40 in total across the range of parameters listed above), simulation, and
plotting scripts in Python are available at
\url{https://github.com/apragsdale/two_locus_selection}.

@@

Mutations at different frequencies are expected to have different signs and
magnitudes of LD. For example, rare mutations show highly positive LD, even for
pairs of neutral mutations \citep{Good2022-ot}. Mutations of different
functional classes (synonymous, missense, and loss-of-function, for example)
are expected to segregate at different average frequencies, and average
selection pressures are observed to be greater within annotated conserved
domains, with stronger purifying selection driving allele frqeuencies to
relatively lower levels among missense and loss-of-function variants. Since
mutations of difference classes and locations have different average
frequencies, this may drive differences in patterns of LD when comparing
between them.

One approach to lessen the impact of differing allele frequencies when
comparing between classes of mutations is to condition on allele frequencies.
From sampling data with observed derived allele counts, it is simplest to
condition on those allele counts \citep[e.g.,\(n_A,n_B=2\),][]{Garcia2021-zn},
and to compute \(\sigma_d^1\) and \(\sigma_d^2\) from pairs of mutations that
satisfy that condition. In the simulations with haploid sample sizes \(n=50\)
(Figures~\ref{fig:toy},~\ref{fig:relate_dom},~\ref{fig:toy_sd2}--\ref{fig:relate_dom_sd2}),
I partitioned allele counts as uncommon (\(n_A, n_B \leq 4\), or \(\hat{f_A},
\hat{f_B} \leq 0.08\)), or common (\(n_A, n_B \geq 5\)). In the
\citet{1000_Genomes_Project_Consortium2015-zq} data, in which most populations
have diploid sample sizes between 80 and 100, I considered \(n_A, n_B \geq 2\),
\(3\leq n_A, n_B \leq 8\), and \(n_A, n_B \geq 9\) (or rare, uncommon, and
common (\(\gtrsim 0.05\))). Functions to compute allele-count-conditioned
statistics from \(\Psi_n\) are packaged within \texttt{moments}.

@@

\subsection{Multinucleotide mutations and positive LD between linked synonymous variants}
\label{sec:mnm}

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{llll}
        \toprule
        Diploid genotype & General model & Simple dominance & Gene-based dominance\\
        \midrule
        $AB$ / $AB$ & $1 + s_{AB/AB}$ & $1 + 2s_A + 2s_B$ & $1 + 2s$\\
        $AB$ / $Ab$ & $1 + s_{AB/Ab}$ & $1 + 2s_A + 2s_Bh_B$ & $1 + 2s$\\
        $AB$ / $aB$ & $1 + s_{AB/aB}$ & $1 + 2s_Ah_A + 2s_B$ & $1 + 2s$\\
        $AB$ / $ab$ & $1 + s_{AB/ab}$ & $1 + 2s_Ah_A + 2s_Bh_B$ & $1 + 2sh$\\
        $Ab$ / $Ab$ & $1 + s_{Ab/Ab}$ & $1 + 2s_A$ & $1 + 2s$\\
        \addlinespace
        $Ab$ / $aB$ & $1 + s_{Ab/aB}$ & $1 + 2s_Ah_A + 2s_Bh_B$ & $1 + 2s$\\
        $Ab$ / $ab$ & $1 + s_{Ab/ab}$ & $1 + 2s_Ah_A$ & $1 + 2sh$\\
        $aB$ / $aB$ & $1 + s_{aB/aB}$ & $1 + 2s_B$ & $1 + 2s$\\
        $aB$ / $ab$ & $1 + s_{aB/ab}$ & $1 + 2s_Bh_B$ & $1 + 2sh$\\
        $ab$ / $ab$ & $1$ & $1$ & $1$\\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{General selection model for diploids and dominance models.}
        Different dominance and interactive effects can be implemented by
        assigning the appropriate values of $s$.
    }
    \label{tab:selmodels}
\end{table}

\begin{table}[ht!]
    \centering
    \begin{tabular}{ll}
        \toprule
        Haplotype & Fitness\\
        \midrule
        $AB$ & $(1 + s_A + s_B)(1+\epsilon)$\\
        $Ab$ & $1 + s_A$\\
        $aB$ & $1 + s_B$\\
        $ab$ & $1$\\
        \bottomrule
    \end{tabular}
    \caption{\textbf{Haploid epistasis model.}}
    \label{tab:epistasis}
\end{table}

\begin{table}[ht!]
    \centering
    \begin{tabular}{lll}
        \toprule
        Code & Description & Region\\
        \midrule
        ESN & Esan in Nigeria & Africa\\
        GWD & Gambian in Western Divisions in the Gambia & Africa\\
        LWK & Luhya in Webuye, Kenya & Africa\\
        MSL & Mende in Sierra Leone & Africa\\
        YRI & Yoruba in Ibadan, Nigeria & Africa\\
        \addlinespace
        CEU & Utah Residents (CEPH) with Northern and Western European Ancestry & Europe\\
        GBR & British in England and Scotland & Europe\\
        FIN & Finnish in Finland & Europe\\
        IBS & Iberian Population in Spain & Europe\\
        TSI & Toscani in Italia & Europe\\
        \addlinespace
        CDX & Chinese Dai in Xishuangbanna, China & East Asia\\
        CHB & Han Chinese in Beijing, China & East Asia\\
        CHS & Southern Han Chinese & East Asia\\
        JPT & Japanese in Tokyo, Japan & East Asia\\
        KHV & Kinh in Ho Chi Minh City, Vietnam & East Asia\\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{
            Thousand Genomes Project population descriptions
            for populations used in this study.
        }
    }
    \label{tab:1kgpops}
\end{table}

\begin{longtable}[t]{lllr}
\caption{\label{tab:tajimasD}
\textbf{Tamija's $D$ for classes of coding mutations.} Here, we partitioned by
synonymous, missense, and nonsense mutationsi, and considered all mutations gene-wide,
mutations falling within annotated conserved elements,
and mutations falling outside of those
elements.}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endfirsthead
\caption[]{\label{tab:tajimasD}Tamija's $D$ for classes of coding mutations.
\textit{(continued)}}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endhead

@@

\begin{table}
    \centering
    \small
    \begin{tabular}{lrrrrrrrrr}
        \toprule
        Class & $h$ & shape & scale & LL &
        $[0,10^{-5})$ & $[10^{-5},10^{-4})$ & $[10^{-4},10^{-3})$ &
        $[10^{-3},10^{-2})$ & $[10^{-2},\infty)$\\
        \midrule
        Missense & 0.0 & 0.093 & 768505 & -678.2 & 0.260 & 0.062 & 0.077 & 0.096 & 0.505\\
         & 0.2 & 0.138 & 6660 & -416.7 & 0.260 & 0.098 & 0.134 & 0.182 & 0.327\\
         & 0.5 & 0.147 & 2117 & -392.0 & 0.282 & 0.114 & 0.159 & 0.214 & 0.231\\
        \addlinespace
        LOF & 0.0 & 0.132 & 99999054 & -248.3 & 0.077 & 0.028 & 0.037 & 0.051 & 0.807\\
         & 0.2 & 0.177 & 477994 & -226.7 & 0.083 & 0.042 & 0.063 & 0.095 & 0.717\\
         & 0.5 & 0.188 & 121419 & -224.2 & 0.092 & 0.050 & 0.077 & 0.119 & 0.662\\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{Distribution of fitness effects inferred for nonsynonymous mutations.}
        DFEs inferred for missense and loss-of-function variants in MSL for varying
        values of $h$. General patterns are consistent across different chosen values
        of $h$, although $h=0$ results in poorer fits for both missense and LOF variants.
        Columns to the right of the log-likelihood (LL) column show proportions of
        new mutations with $|s|$ in each given bin.
    }
    \label{tab:msldfe}
\end{table}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/jackknife}
    \caption{
        \textbf{Accuracy of the jackknife approximation and runtime.}
        Small sample sizes can lead to large error in the moment-closure
        approximation for larger recombination distances or selection
        coefficients. Generally, the jackknife approximation breaks down for
        recombination rates greater than \(\rho\gtrsim50\), depending on
        the sample size. While increasing
        sample size leads to more accurate solutions, it comes at the cost of
        both increased runtime and memory usage. Most analyses in this paper
        used sample sizes between 40 and 80.
    }
    \label{fig:jackknife}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../simulations/discrete/plots/comp_Ne_5000_n_50_r_0.0_s_0.0_e_0.0.pdf}
    \caption{
        \textbf{Comparison between numerics and simulations for fully linked neutral loci.}
        With no recombination (\(r=0\)) and no selection, the moments solution is
        exact, and computing \(sigma_d^2\) from \(\Psi_n\) gives 5/11, the analytic
        solution from \citet{Ohta1969-ie}.
        Discrepancies between numerical expectations and the simulations are
        due to simulation noise, with Anscombe Poisson residuals roughly normally
        distributed around zero \citep{Pierce1986-um}.
        In the bar plots, ``E'' (blue) are expectations from \texttt{moments} and
        ``D'' are simulated data.
    }
    \label{fig:validation1}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../simulations/discrete/plots/comp_Ne_5000_n_50_r_0.001_s_0.002_e_0.pdf}
    \caption{
        \textbf{Comparison to simulations for loosely linked additively selected loci.}
        \(Ne=5000\), \(r=0.001\), and \(s=-0.002\) (\(\rho=20\) and \(\gamma=-20\)
        at both loci), with no epistatic interactions (\(\epsilon=0\)).
        E: \texttt{moments} expectations, D: simulated data.
    }
    \label{fig:validation2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../simulations/discrete/plots/comp_Ne_5000_n_50_r_5e-05_s_0.0002_e_-0.5.pdf}
    \caption{
        \textbf{Comparison to simulations with antagonistic epistasis.}
        \(Ne=5000\), \(r=0.00005\), \(s=-0.0002\) (\(\rho=1\) and \(\gamma=-2\)
        at both loci), and \(\epsilon=-0.5\).
        E: \texttt{moments} expectations, D: simulated data.
    }
    \label{fig:validation3}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../simulations/discrete/plots/comp_Ne_5000_n_50_r_5e-05_s_0.0002_e_-0.5.pdf}
    \caption{
        \textbf{Comparison to simulations with synergistic epistasis.}
        \(Ne=5000\), \(r=0.00005\), \(s=-0.0002\) (\(\rho=1\) and \(\gamma=-2\)
        at both loci), and \(\epsilon=0.5\).
        E: \texttt{moments} expectations, D: simulated data.
    }
    \label{fig:validation4}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../simulations/discrete/plots/comp_Ne_5000_n_50_r_5e-06_s_0.0002_h_0.1.pdf}
    \caption{
        \textbf{Comparison to simulations with site-wise dominance.}
        \(Ne=5000\), \(r=5\times10^{-6}\), and \(s=-0.0002\) (\(\rho=0.1\) and
        \(\gamma=-2\)), with \(h=0.1\) at both loci and no
        epistatic interactions (\(\epsilon=0\)).
        E: \texttt{moments} expectations, D: simulated data.
        Additional comparisons to simulated data are availabe at
        \texttt{https://github.com/apragsdale/two\_locus\_selection}.
    }
    \label{fig:validation5}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with slightly deleterious mutations.}
        Using an individual based forward simulator \citep{Thornton2019-qc}, I
        simulated many replicates of 1Mb regions with high mutation rates of
        selected mutations (see Section~\ref{sec:bgs} for details). Here,
        many linked additive weakly selected mutations slightly distort the
        SFS for both selected and neutral mutations. \(\sigma_d^2\) is reduced
        relative to two-locus expectations within background selection, and
        Hill-Robertson interference effects are slightly diminished with
        \(\sigma_d^1\) not as strongly negative as expected.
    }
    \label{fig:bgs1}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with moderately deleterious mutations.}
        In the case of many linked strongly selected additive mutations, neither
        selected nor neutral statistics deviate dramatically from two-locus expectations.
        The SFS from the simulated data matches expectations, with a slight excess
        of selected mutations. LD is also largely unaffected, with \(\sigma_d^2\)
        matching expectations and \(\sigma_d^1\) fluctuating near zero as expected.
    }
    \label{fig:bgs2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.0_n_50}
    \caption{
        \textbf{Associative overdominance due to slightly deleterious recessive mutations.}
        The presence of many linked weakly selected recessive mutations can cause
        measues of diversity to have large increases over single- and two-locus
        expectations. The SFS for both neutral and selected mutations both show
        a very large excess of common variants (note the log scale), and \(\sigma_d^2\)
        are also increased over two-locus expectations for both classes of mutations.
        Still, \(\sigma_d^1\) remains 0 for neutral mutations and is not as strongly
        negative for pairs of selected mutations.
    }
    \label{fig:bgs3}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.1_n_80}
    \caption{
        \textbf{Background selection with moderately deleterious recessive mutations.}
        More strongly deleterious recessive mutations do not show as dramatic of
        deviations from expectations as the case of weakly deleterious recessive
        mutations. The neutral SFS is reduced, and we observe more selected mutations
        at higher frequencies than expected. LD is also only weakly distorted from
        two-locus expectations. Notably, we see positive LD (matching between
        simulations and expectations) for strongly deleterious recessive mutations,
        as predicted in \citet{Roze2021-cf}.
    }
    \label{fig:bgs4}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.sd2}
    \caption{
        \textbf{Time series of squared LD under bottleneck 
        and expansion histories with epistasis.}
        Bottlenecks increase \(\sigma_d^2\) (squared LD),
        while expansions tend to decrease squared LD.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance}
    \caption{
        \textbf{Time series of signed LD under bottleneck 
        and expansion histories with dominance.}
        Gene-based dominance, unlike site-wise dominance effects, can cause
        large positive signed LD, similar to compensatory mutations or
        antagonistic epistasis. In general, common variants are more
        stable under population size changes than rare variants.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_dom}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance.sd2}
    \caption{
        \textbf{Time series of squared LD under bottleneck 
        and expansion histories with dominance.}
        Similar to the cases with epistasis, bottlenecks tend to increase
        \(\sigma_d^2\) and expansions decrease squared LD.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_dom_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU}
    \caption{
        \textbf{Time series of signed LD under inferred models of human population
        size history with with epistasis.}
        Similar to the instantaneous size changes in the toy models
        (Figure~\ref{fig:toy}), bottlenecks tend to make signed LD more extreme.
        However, in the case of strong selection and antagonistic epistasis,
        the CEU \(\sigma_d^1\) trajector decreases relative to the YRI
        trajectory. Again, statistics for rare variants fluctuate more rapidly
        than common variants.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.sd2}
    \caption{
        \textbf{Time series of squared LD under inferred models of human population
        size history with with epistasis.}
        Size reductions, such as bottlenecks in the recent past of Eurasian populations,
        increase squared LD.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.dominance.sd2}
    \caption{
        \textbf{Time series of squared LD under inferred models of human population
        size history with with dominance.}
        Again, bottlenecks increase squared LD relative to populations that did
        not experience such strong size reductions in their history.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate_dom_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_gene_wide}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        in the same gene.}
        When taking gene-wide averages of LD, there are no obvious differences
        between the LD decay curves of synonymous and missense mutations.
        Each panel contains five LD decay curves for each mutation class, one for
        each \citet{1000_Genomes_Project_Consortium2015-zq} population in the
        continental groupings. Shared demographic history leads to strong
        correlation in statistics between populations in the same continental
        groups.
    }
    \label{fig:LDgene}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_in_domain}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        that fall inside the same domain.}
        At short distances between SNPs that both fall within sthe same conserved
        annotated domain, missense mtuations show consistently
        larger signed LD than synonymous variants. This may be caused by
        antagonistic epistasis or compensatory effects between pairs of
        nearby missense mutations.
    }
    \label{fig:LDwithin}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_outside_domains}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        outside of domains.}
        Outside of conserved domains,
        missense mutations have reduced signed LD at short distances
        compared to synonymous mutations,
        opposite the pattern observed within domains (Figure~\ref{fig:LDwithin}).
        This pattern is consistent with Hill-Robertson interference between
        missense mutations outside of conserved elements.
    }
    \label{fig:LDoutside}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_afr}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes African group.
        We observe increased LD between missense mutations within domains,
        but decrased LD outside of domains at matched distances. This pattern
        holds when conditioning on non-rare variation (\(n_A, n_B \geq 9\),
        MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsAFR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eur}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes European group.
        We again observe increased LD between missense mutations within domains,
        but decrased LD outside of domains at matched distances. And again,
        this pattern holds when conditioning on non-rare variation
        (\(n_A, n_B \geq 9\), MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsEUR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eas}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes East Asian group.
        As with the African and European data,
        wobserve increased LD between missense mutations within domains,
        but decrased LD outside of domains at matched distances. And again,
        this pattern holds when conditioning on non-rare variation
        (\(n_A, n_B \geq 9\), MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsEAS}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../figures/data_within_between}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations both within the same
        domain and in differing domains.}
        In each panel (one for each population), the left two data points show signed
        LD between mutations that fall within the same domain, and the right two
        data points show signed LD between mutations falling within domains but
        in \emph{different} domains within the same protein-coding gene. Distances
        are not matched between these data classes (as between-domain pairs are
        typically more distant than within-domain pairs). Increased LD between
        missense mutations is only observed for pairs of mutations within the same
        domain, but not between domains.
    }
    \label{fig:between_domains}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/mutations_in_genes}
    \caption{
        \textbf{Correlation between numbers of missense and synonymous mutations
        within each gene.}
        Human genes have different sizes and sequence content, so they may differ in the
        numbers of missense and synonymous mutations observed within them.
        The number of missense mutations within a gene is predicted well by
        the number of synonymous mutations (\(r=0.86\)).
    }
    \label{fig:mutCorrGene}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/mutations_in_domains}
    \caption{
        \textbf{Correlation between numbers of missense and synonymous mutations
        within each annotated domain.}
        Domains are typically much smaller than the average gene-size, and while
        the number of synonymous mutations within a domain is moderately predictive
        of the number of missense mutations within the same domain (\(r=0.57\)),
        this correlation is weaker than the correlation in gene-wide mutation
        counts (Figure~\ref{fig:mutCorrGene}).
    }
    \label{fig:mutCorrDomain}
\end{figure}
